---
title: "PA1_template.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pressure, echo=FALSE}
plot(pressure)
#Code for reading in the dataset and/or processing the data
echo = TRUE
getwd()


full_data <- read.csv("activity.csv", header = TRUE)

# Histogram of the total number of steps taken each day
steps_each_day <- aggregate(steps ~ date, full_data, sum)
hist(steps_each_day$steps, breaks = 20,
     main = paste("Total Number of Steps Taken Each Day")
     , xlab = "Number of Steps")

#Mean and median number of steps taken each day
mean_steps <- mean(steps_each_day$steps)
median_steps <- median(steps_each_day$steps)
#Calculate and report the mean and median of the total number of steps taken per day
sprintf("Mean of the steps taken each day %.3f", mean_steps)
sprintf("Median of the steps taken each day %.3f", median_steps)


#Time series plot of the average number of steps taken
avg_steps_across_all_days <- aggregate(steps ~ interval, full_data, mean)
plot(avg_steps_across_all_days$interval, avg_steps_across_all_days$steps, type = "l", xlab = "Interval"
     , ylab = "Average Steps", main = "Average Number of Steps per Day by Interval")

#The 5-minute interval that, on average, contains the maximum number of steps
max_step_interval <- avg_steps_across_all_days[which.max(avg_steps_across_all_days$steps),1]
sprintf("Maximum number of steps were taken in interval %d", max_step_interval)

#Code to describe and show a strategy for imputing missing data
sprintf("Total number of rows with missing values is %d", sum(is.na(full_data)))


library(Hmisc)
## Copy full data into a new data set
full_data_imputed <- full_data
full_data_imputed$weekday <- weekdays(as.Date(full_data$date))

## Replacing missing values under steps column with mean of the same column
full_data_imputed$steps <- ave(full_data$steps, full_data_imputed$weekday,FUN=function(x) ifelse(is.na(x), mean(x,na.rm=TRUE),x))

steps_each_day_imputed <- aggregate(steps ~ date, full_data_imputed, sum)

#Histogram of the total number of steps taken each day after missing values are imputed
hist(steps_each_day_imputed$steps, breaks = 20, col = 'red' 
     ,main = paste("Total Number of Steps Taken Each Day"), xlab = "Number of Steps", ylim = c(0,20))


hist(steps_each_day$steps, breaks = 20, col = 'red', main = paste("Total Number of Steps Taken Each Day"), 
     xlab = "Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"),lwd = 10)

mean_steps_imputed <- mean(steps_each_day_imputed$steps)
median_steps_imputed <- median(steps_each_day_imputed$steps)

comparison_table <- matrix(c(mean_steps, mean_steps_imputed, median_steps, median_steps_imputed), ncol = 2, byrow = TRUE)

#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
colnames(comparison_table) <- c("Original", "Imputed")
rownames(comparison_table) <- c("mean", "median")
full_data_imputed$weekend_factor <- ifelse(weekdays(as.Date(full_data_imputed$date)) %in%
                                             c("Saturday", "Sunday"), "weekend", "weekday")
panel_plot <- aggregate(steps ~ interval + weekend_factor, data = full_data_imputed, mean)
library(ggplot2)
ggplot(panel_plot, aes(interval, steps)) + geom_line() + facet_grid(weekend_factor ~ .) 
+ xlab("5 min Interval") + ylab("Average Steps") + ggtitle("Weekdays vs. Weekends Average Steps Pattern") + 
  theme(plot.title = element_text(hjust = 0.5))